GravExplain: Continuous gravitational wave searches in a table-top experiment

### 2020-02-12 paper/paper_main.tex
74th day of Summer

% cover page
\section*{Project contributions}
% contributions clarify what’s to be assessed
Hannah Middleton: introduction of paper, design and making of interferometer

James W. Gardner: all other paper sections, design and making of photodiode circuit, and all of the source code

\jam{do we need to repeat those people in the Acknowledgements?}

@@

\author{Hannah Middleton}
\email{hannah.middleton@unimelb.edu.au}
\author{Andrew Melatos}
\email{amelatos@unimelb.edu.au}
\affiliation{School of Physics, University of Melbourne, Parkville, VIC, 3010, Australia}
\affiliation{OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery}

@@

This equation also means that the intensity changes linearly with a small change in separation, as in Equation~\ref{eq:intensity_change} (see Appendix~\ref{app:intensity_derivation} again for the derivation).

\begin{equation}
\label{eq:intensity_change}
	\delta I = \frac{- 8 \pi A^2 \cos{\theta} \sin{\phi}}{\lambda}\, \delta d
\end{equation}

@@

The entire circuit diagram is detailed and shown in Appendix~\ref{app:circuit_diagram}. The updated schematic is shown in Figure~\ref{fig:ifo_schematic_podo}.

@@

@@

A full model of the transfer function would detail the system starting from the voltage sent to the speaker and ending with the value recorded by the Pi, including any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading. Although we have the separation-intensity relation as linear for small motions, the other components of this full model are yet to be determined.

@@

\section{Open-source code}

\jam{update this to the current README.txt}

This project is implemented in python3\cite{python} inside of jupyter notebooks\cite{jupyter}\cite{ipython} and uses the packages of numpy\cite{numpy}, scipy\cite{scipy}, matplotlib\cite{matplotlib}, tqdm\cite{tqdm}, and logmmse\cite{logmmse}.

Current build found at:
\url{https://github.com/daccordeon/gravexplain}

\begin{verbatim}
Directory structure:
gravexplain/
    source/
        [py executable script versions of all .ipynb notebooks]
        raspberry_pi_photodiode_adc.py
            (reads photodiode signal through ADC by SPI pins, saves to .csv)
    paper/
        paper_main.tex
        ifoDemoBib.bib
        myunsrt.bst

    mock_continuous_signal.ipynb
        (simulates applying the Viterbi algorithm to a noisy signal)
    experiment_viterbi.ipynb
        (applies the Viterbi algorithm to an .mp4 from the webcam method)
    experiment_photodiode.ipynb
        (filters signal from optical microphone produced by raspberry_pi_photodiode_adc.py,
        also applies the Viterbi algorithm)

    tracker_time_series.ipynb
        (defunct; saves signal and spectrum plots from a .csv file)
    open_cv.ipynb
        (defunct; various functions produced while learning the openCV library)

    .gitignore
    LICENSE
    README.txt
\end{verbatim}

@@

Then take derivatives and find that the intensity changes linearly for small changes in separation.

\begin{align}    
    \frac{\delta I}{\delta d} &= \frac{\delta I}{\delta \phi}\; \frac{\delta \phi}{\delta d}\\
    \frac{dI}{d\phi} &= - 2 A^2 \sin{\phi}\\
    \delta\phi &= \frac{4 \pi \cos{\theta}}{\lambda}\, \delta d\\
    \therefore \delta I &= \frac{- 8 \pi A^2 \cos{\theta} \sin{\phi}}{\lambda}\, \delta d
\end{align}

@@

\begin{acknowledgments}
The authors are grateful to Robin Evans, Bill Moran, Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project.

>>>>

% cover page
\section*{Project contributions}
% contributions clarify what’s to be assessed
James W. Gardner: all paper sections other than the introduction, design and making of the photodiode circuit, and all of the source code at \url{https://github.com/daccordeon/gravexplain}

Hannah Middleton: introduction of paper, design and making of the interferometer

Andrew Melatos, Robin Evans, William Moran: vital advice and comments throughout the project

@@

\author{Hannah Middleton$^{(i)}$}
\email{hannah.middleton@unimelb.edu.au}
\author{Andrew Melatos$^{(i)}$}
\email{amelatos@unimelb.edu.au}
\author{Robin Evans$^{(ii)}$}
\email{robinje@unimelb.edu.au}
\author{William Moran$^{(ii)}$}
\email{wmoran@unimelb.edu.au}
\affiliation{School of Physics$^{(i)}$ and Department of Electrical and Electronic Engineering$^{(ii)}$, University of Melbourne, Parkville, VIC, 3010, Australia}
\affiliation{OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery}

@@

This equation means that the intensity changes non-linearly with a small change in separation, as in Equation~\ref{eq:intensity_change} (see Appendix~\ref{app:intensity_derivation} again for the derivation). This makes the problem of extracting the original signal more difficult.

\begin{equation}
\label{eq:intensity_change}
	\frac{\delta I}{\delta d} = \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
\end{equation}

@@

The entire circuit diagram is detailed and shown in Appendix~\ref{app:circuit_diagram}. The updated schematic is shown in Figure~\ref{fig:ifo_schematic_podo} and a picture taken of the experiment, showing the interferometer, the circuit, and the Pi, is shown in Figure~\ref{fig:setup_pic2}.

@@

\begin{figure}
	\includegraphics[width=0.8\textwidth]{figures/setup_pic2.pdf}
	\caption{Picture taken of the optical microphone set-up, showing the Michelson interferometer, the photodiode (in an insert), the circuit, and the Raspberry Pi.}
	\label{fig:setup_pic2}
\end{figure}

@@

A full model of the transfer function would detail the system starting from the voltage sent to the speaker and ending with the value recorded by the Pi, including any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading. Although we have the separation-intensity relation as non-linear for small motions, the other components of this full model are yet to be determined.

@@

\section{Open-source code}
This project is implemented in python3\cite{python} inside of jupyter notebooks\cite{jupyter}\cite{ipython} and uses the packages of numpy\cite{numpy}, scipy\cite{scipy}, matplotlib\cite{matplotlib}, tqdm\cite{tqdm}, and logmmse\cite{logmmse}.

The current build can be found at:
\url{https://github.com/daccordeon/gravexplain}
% \begin{verbatim}
% README
% \end{verbatim}

@@

Then take derivatives and find that the intensity change is non-linear.

\begin{align}    
    \frac{\delta I}{\delta d} &= \frac{\delta I}{\delta \phi}\; \frac{\delta \phi}{\delta d}\\
    \frac{\delta I}{\delta\phi} &= - 2 A^2 \sin{\phi}\\
    \frac{\delta\phi}{\delta d} &= \frac{4 \pi \cos{\theta}}{\lambda}\\
    \therefore \frac{\delta I}{\delta d} &= \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
\end{align}

@@

\begin{acknowledgments}
The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project.
